
For the first dataset, we are looking at houses are are only within the [NAmes, Edwards, BrkSide neighborhoods] and are looking at how 

SalePrice ~ SqFt (GrLIvArea) + Neighborhood 

and if SalesPrice depends on neighborhood. 

```{r echo=FALSE}
library(tidyverse)
library(plotly)

housing = read.csv(file = "train.csv")
a1_set = housing %>% select(SalePrice, Neighborhood, GrLivArea) %>% filter( Neighborhood == c("NAmes", "Edwards", "BrkSide"))
```

Initializing any required packages and calling in the data with some light filtering, we can take a look at the dataset and visualize what the datalooks like with some cursory visuals. 

```{r}

a1_set %>% ggplot(aes(x = (GrLivArea), y = SalePrice/1000, color = Neighborhood)) + geom_point() + stat_smooth(method = "lm")

a1_set %>% ggplot(aes(x = SalePrice/10000, fill = Neighborhood)) + geom_histogram(bins =20)
```

We can see in the dataset that there is a difference in slope/intercept values per group 

```{r}
#Comparison of Means/Coefficients to signify a difference in sample groups from different neighborhoods.
n_means <- aov(SalePrice ~ Neighborhood, data = a1_set)
mean_diff <- TukeyHSD(n_means, conf.level = 0.95)
mean_diff
fit <- lm(SalePrice ~ GrLivArea + Neighborhood + GrLivArea*Neighborhood, data = a1_set)
```

